---
title: "Getting Started with CEClust"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with CEClust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

`CEClust` explores composite entropy clustering over one value of `lambda`, a
lambda path, or a full `(C, lambda)` grid.

The usual workflow is:

1. fit a grid with `CECfitBoundGrid()`;
2. inspect the phase diagram with `CECplotGrid()`;
3. inspect selected partitions with `CECplotPartition()`;
4. summarise diagnostics with `CECsummariseGrid()`;
5. optionally open the Shiny explorer with `CECexplore()`.

# A small iris grid

This example uses only the four quantitative iris variables. The species labels
are not used in the clustering step.

```{r}
library(CEClust)

Z_iris <- iris[, 1:4]

iris_grid <- CECfitBoundGrid(
  Z = Z_iris,
  dataset_name = "iris_vignette",
  algo_seed = 202606,
  lambda_grid = seq(0.6, 1.4, by = 0.4),
  C_grid = c(1.5, 2.5),
  r0 = 6,
  k0 = 2,
  B = 2,
  Nshots_fresh = 1,
  Nshots_warm = 1,
  Nloop = 10,
  familyType = "gaussVector",
  n_cores = 1,
  compact_results = TRUE
)

CECplotGrid(iris_grid)
```

# Inspect one partition

```{r}
CECplotPartition(iris_grid)
```

# Summarise diagnostics

```{r}
iris_summary <- CECsummariseGrid(iris_grid)
head(iris_summary)
```

For larger grids, use:

```{r, eval = FALSE}
CECexplore(iris_grid)
```
