---
title: "Getting Started with CEClust"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with CEClust}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

`CEClust` is built around two complementary tasks:

1. fit a composite entropy clustering model for a fixed value of `lambda`;
2. explore how the selected partition evolves over a grid of `lambda` values.

The second task is often the most informative in practice, because it reveals
stable and unstable regions instead of forcing a single regularisation choice
too early.

# Runtime configuration

Before fitting a model, it is useful to choose a runtime profile.

```{r}
library(CEClust)

runtime_base <- CECconfigure_runtime("base")
runtime_base
```

The `"base"` profile forces a conservative single-core pure-R execution.
When available, the `"auto"` profile can activate the optional compiled backend
and suggest a larger number of worker processes:

```{r}
runtime_auto <- CECconfigure_runtime("auto", max_cores = 4)
runtime_auto
```

The returned `n_cores` value can be passed directly to
`CECfitLambdaGrid()`.

# Example data

For a lightweight mixed-data example:

```{r}
Z <- simulate_multidim_benchmark_data(
  n = 120,
  p_num = 3,
  p_fac = 2,
  K = 3,
  seed = 1
)

str(Z)
```

# Step 1: fit one model for one lambda

`CECclassif()` is the fixed-lambda entry point.

```{r}
fit <- CECclassif(
  Z = Z,
  lambda = 0.8,
  C = 10,
  r0 = 6,
  Nshots = 4,
  Nloop = 20,
  familyType = "gaussAndDiscreteVector"
)

fit$REO
table(fit$clusters)
```

This is useful when `lambda` is already fixed by a methodological decision or
when you want to inspect one candidate value quickly.

# Step 2: diagnose a lambda grid

The main workflow follows linked paths over a grid of `lambda` values:

```{r}
lambda_diag <- CECfitLambdaGrid(
  Z = Z,
  lambda_grid = seq(0.1, 2.0, by = 0.2),
  k0 = 10,
  B = 10,
  C = 10,
  r0 = 10,
  Nshots_fresh = 1,
  Nshots_warm = 1,
  Nloop = 15,
  familyType = "gaussAndDiscreteVector",
  seed = 1,
  silent = TRUE,
  verbose = FALSE,
  n_cores = runtime_base$n_cores,
  checkpoint_dir = FALSE,
  auto_checkpoint = FALSE,
  show_progress = FALSE
)

lambda_diag$summary
```

The `summary` table contains, for each `lambda`:

- stability summaries on the original data;
- bootstrap stability summaries;
- criterion summaries (`H`);
- information about the best run origin.

# Step 3: smooth the summary and identify stable regions

```{r}
lambda_diag <- CECaddSmoothedDiagnostics(lambda_diag, k = 3)

stable <- CECidentifyStableLambdas(
  lambda_diag,
  rule = "ratio",
  ratio_threshold = 0.8, 
  stabB_threshold = 0.8,
  min_consecutive = 1
)

stable$lambda_min
stable$summary
```

`stable$lambda_min` gives the first lambda entering a stable regime under the
selected rule, but the full `stable$summary` table is usually more informative
because it shows whether stability persists over an interval.

When you also want to exclude lambdas for which too much of the sample lives
in clusters saturating the density bound `C`, first extract one representative
partition per lambda and then pass it back to `CECidentifyStableLambdas()`:

```{r}
best_parts <- CECextractBestPartitions(
  lambda_diag,
  source = "fixed",
  criterion = "projected_H",
  Z = Z
)

stable_sat <- CECidentifyStableLambdas(
  lambda_diag,
  rule = "ratio",
  ratio_threshold = 0.8,
  stabB_threshold = 0.8,
  sat_threshold = 0.8,
  best_partitions_obj = best_parts,
  min_consecutive = 1
)

stable_sat$summary
```

# Step 4: extract one partition per lambda

```{r}
best_parts <- CECextractBestPartitions(
  lambda_diag,
  source = "fixed",
  criterion = "projected_H",
  Z = Z
)

best_parts$summary
```

The extracted object is a compact summary of the lambda trajectory. It is often
the easiest object to pass to downstream plotting or diagnostic helpers.

# Step 5: inspect coherence and change points

```{r}
coherence <- CECcheckBestPartitionCoherence(best_parts, Z)
changes <- CECdetectPartitionChanges(best_parts)

coherence$all_coherent
changes$change_lambdas
```

If the trajectory is locally inconsistent, a repair pass can be attempted:

```{r}
best_parts_repaired <- CECrepairBestPartitionTrajectory(best_parts, Z, max_iter = 20)
best_parts_repaired$n_repairs
best_parts_repaired$converged
```

# Static plots

The simplest overview is the lambda-grid diagnostic plot:

```{r lambda-grid-plot, fig.width=9, fig.height=7}
plotCECdiagnoseLambdaGrid(lambda_diag)
```

For multivariate numeric data, a PCA view can be built from extracted best
partitions. The `iris` data set is a convenient example because it contains
four numeric variables.

```{r}
Z_iris <- iris[, -5]

diag_iris <- CECfitLambdaGrid(
  Z = Z_iris,
  lambda_grid = seq(0.01, 2, by = 0.1),
  k0 = 10,
  B = 10,
  C = 10,
  r0 = 20,
  Nshots_fresh = 10,
  Nshots_warm = 10,
  Nloop = 20,
  familyType = "gaussVector",
  seed = 1,
  silent = TRUE,
  verbose = FALSE,
  n_cores = 1L,
  checkpoint_dir = FALSE,
  auto_checkpoint = FALSE,
  show_progress = FALSE
)

best_iris <- CECextractBestPartitions(diag_iris, source = "fixed", Z = Z_iris)
path_df <- build_partition_path_df(Z_iris, best_iris)
plot_partition_path_pca(path_df)
```

# Interactive exploration

If `shiny` and `ggplot2` are installed, you can launch the interactive explorer
from either a `lambda_diag` object or an already extracted `best_parts` object:

```{r eval = FALSE}
plotCECdiagnoseInteractive(best_parts, Z)
```

For one-dimensional numeric data, the application switches automatically to a
histogram display. For higher-dimensional numeric data, the user can switch
between raw variable pairs and a PCA representation.

# Remarks

- Use `CECclassif()` when you want one clustering solution for one `lambda`.
- Use `CECfitLambdaGrid()` when you want to study robustness
  across a grid.
- Use `CECextractBestPartitions()` to obtain a compact trajectory that is easy
  to plot or compare.
- Use `CECidentifyStableLambdas()` as a decision aid, not as a substitute for
  methodological judgement.
