---
title: "Minimal CEClust API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimal CEClust API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The public CEClust API is organised around a small number of verbs:

- `CECclassif()` fits one model for one `lambda`;
- `CECfitLambdaGrid()` scans a lambda grid for one fixed `C`;
- `CECfitBoundGrid()` scans a `C x lambda` grid;
- `CECfitPreset()` launches named workflows such as `"iris"`,
  `"breast_cancer"`, and `"gaussian_1d"`;
- `CECselectStableLambdas()` and `CECsummariseGrid()` provide diagnostics;
- `CECplotGrid()`, `CECplotPartition()`, and `CECplotPath()` provide static
  visual checks;
- `CECexplore()` launches the Shiny explorer.

# Preset workflows

Preset workflows are intended for examples and reproducible checks.

```{r eval = FALSE}
library(CEClust)

iris_grid <- CECfitPreset(
  "iris",
  lambda_grid = seq(0.4, 1.0, by = 0.2),
  C_grid = c(1.5, 2.0),
  k0 = 2,
  B = 2,
  Nshots_fresh = 1,
  Nshots_warm = 1,
  Nloop = 10,
  n_cores = 1,
  save_results = FALSE
)
```

The returned object contains one row per `C` and `lambda` pair:

```{r eval = FALSE}
head(CECsummariseGrid(iris_grid))
```

# Diagnostics

Stable lambda selection combines the existing stability, RSI, and saturation
criteria:

```{r eval = FALSE}
stable <- CECselectStableLambdas(
  iris_grid,
  C = 2.0,
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.8,
  sat_threshold = 0.05
)

stable$kept_lambdas
```

# Plots and Shiny

The static plot layer uses the same grid object:

```{r eval = FALSE}
CECplotGrid(iris_grid)
CECplotPartition(iris_grid)
CECplotPath(iris_grid)
```

The Shiny explorer is kept as the main interactive interface:

```{r eval = FALSE}
CECexplore(iris_grid)
```

# Direct C by lambda grids

For custom data, use `CECfitBoundGrid()`:

```{r eval = FALSE}
Z <- simulate_multidim_benchmark_data(n = 120, p_num = 3, p_fac = 0, seed = 1)

grid <- CECfitBoundGrid(
  Z = Z,
  lambda_grid = seq(0.4, 1.0, by = 0.2),
  C_grid = c(1, 2),
  familyType = "gaussVector",
  k0 = 2,
  B = 2,
  Nshots_fresh = 1,
  Nshots_warm = 1,
  Nloop = 10,
  n_cores = 1
)
```

