---
title: "Minimal CEClust API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimal CEClust API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

The minimal CEClust API is organised around a small number of verbs:

- `CECclassif()` fits one model for one pair `(C, lambda)`;
- `CECfitLambdaGrid()` scans a lambda grid for one fixed `C`;
- `CECfitBoundGrid()` scans a full `(C, lambda)` grid;
- `CECfitPreset()` launches named workflows such as `"gaussian_1d"`,
  `"iris"`, and `"breast_cancer"`;
- `CECplotGrid()`, `CECplotPath()`, and `CECplotPartition()` are the first
  inspection tools;
- `CECexplore()` launches the Shiny explorer;
- `CECsummariseGrid()` and `CECselectStableLambdas()` provide diagnostic
  tables once the plots have been inspected.

By default, grid results stay in memory. The public wrappers do not write
checkpoint or result files unless this is requested explicitly.

# Visual-first workflow

A CEClust analysis should usually begin with a grid and plots. The diagnostic
numbers are useful, but they are easier to understand after seeing the
structure of the path.

```{r}
library(CEClust)

article_grid <- CECfitPreset(
  "gaussian_1d",
  n = 100,
  data_seed = 13,
  algo_seed = 2026035,
  lambda_grid = seq(0.1, 1.8, by = 0.1),
  C_grid = c(0.5, 1, 2, 3, 3.5, 4, 4.5, 5, 6, 8),
  r0 = 10,
  k0 = 20,
  B = 20,
  Nshots_fresh = 10,
  Nshots_warm = 10,
  Nloop = 100,
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.8,
  sat_threshold = 0.05
)
```

The first plot is the map of the `(C, lambda)` grid:

```{r}
CECplotGrid(article_grid)
```

For one-dimensional data, the partition path is often more informative than
the diagnostic table:

```{r}
CECplotPath(
  article_grid,
  C = 1,
  type = "partitions",
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.8,
  sat_threshold = 0.05,
  saturation_mark = "hatch"
)
```

Then inspect a selected cell:

```{r}
CECplotPartition(article_grid, C = 1, lambda = 1)
```

# One fixed classification

The simple `CECclassif()` entry point is still useful when `C` and `lambda` are
chosen in advance, or after a grid suggests a reasonable cell.

```{r}
fit_1_1 <- CECclassif(
  Z = article_grid$Z,
  C = 1,
  lambda = 1,
  r0 = 10,
  Nshots = 50,
  Nloop = 100,
  familyType = "gaussUniv"
)

table(fit_1_1$clusters)
fit_1_1$REO
```

# Diagnostics

After the visual checks, summarise the grid and select stable lambdas:

```{r}
head(CECsummariseGrid(article_grid))

stable_C1 <- CECselectStableLambdas(
  article_grid,
  C = 1,
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.8,
  sat_threshold = 0.05
)

stable_C1$kept_lambdas
stable_C1$summary
```

# Shiny explorer

The Shiny interface is kept as the main interactive workflow:

```{r}
CECexplore(article_grid)
```

# Iris presets

Run iris without using `Species` in the clustering variables:

```{r}
iris_no_species <- CECfitPreset(
  "iris",
  include_species_as_feature = FALSE,
  quantitative_representation = "raw",
  r0 = 10,
  k0 = 20,
  B = 20,
  Nshots_fresh = 10,
  Nshots_warm = 10,
  Nloop = 100
)

CECplotGrid(iris_no_species)
CECplotPartition(iris_no_species)
CECexplore(iris_no_species)
```

Run iris with `Species` included as a qualitative feature:

```{r}
iris_with_species <- CECfitPreset(
  "iris",
  include_species_as_feature = TRUE,
  quantitative_representation = "raw",
  r0 = 10,
  k0 = 20,
  B = 20,
  Nshots_fresh = 10,
  Nshots_warm = 10,
  Nloop = 100
)

CECplotGrid(iris_with_species)
CECplotPartition(iris_with_species, shape_var = "Species")
CECexplore(iris_with_species)
```

# Uniform stress test

For a uniform sample on `[0, 1]`, use a lambda path at `C = 10` and mark
values that do not satisfy `RSI > 0.9`, `stab_algo > 0.8`, and `sat = 0`.

```{r}
set.seed(13)
Z_uniform <- runif(1000)

uniform_grid <- CECfitBoundGrid(
  Z = Z_uniform,
  lambda_grid = seq(0.05, 2, by = 0.05),
  C_grid = 10,
  r0 = 50,
  k0 = 20,
  B = 20,
  Nshots_fresh = 20,
  Nshots_warm = 20,
  Nloop = 100,
  familyType = "gaussUniv",
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.9,
  sat_threshold = 0
)

CECplotPath(
  uniform_grid,
  C = 10,
  type = "partitions",
  stab_algo_threshold = 0.8,
  rsi_threshold = 0.9,
  sat_threshold = 0,
  saturation_mark = "hatch"
)
```
